### 2021-06-07 AggregateCounts.Rmd
changed folder and got aggregates files import

---
title: "AggregateCounts"
author: "Juan P Barraza"
date: "6/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(conflicted)
library(tidyverse)
filter <- dplyr::filter

conditionlist= list('wt','pqsL', 'phz', 'mono')

ReadAggregateFile(fileList[1], "Pa", "1", "wt")
GetAggregates()

datos <-read.csv(fileList[1], header = F)
colnames(datos)<-c('Size')
datos1<-datos %>% filter(!Size<0.5) #first filter the size

ReadAggregateFile<-function(file, species, timepoint, condition){#make a dataframe of the aggregate size list and add experimental parameters
  datos <-read.csv(file, header = F)
  colnames(datos)<-c('Size')
  datos<-datos %>% filter(!Size<0.5) #first filter the size
  datos<-datos %>% mutate(Species = species)#add parameters
  datos<-datos %>% mutate(TimePoint = timepoint)
  datos<-datos %>% mutate(Condition = condition)
  return(datos)
}

GetAggregates<-function(){#import all experiments at once
  folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (condition in 1:length(conditionlist)){
    currentCondition<-fileList[grep(conditionlist[condition],fileList)] #find files for each condition
    print(currentCondition)
    for (species in 1:3){
      currentTimePoint<-currentPosition[grep(paste0("t_",GetNumber(timepoint),"p_"),currentPosition)]
      print(currentTimePoint)
      #growthDf<-rbind(growthDf,ReadAggregateFile(currentTimePoint, position, timepoint))
    }
  }
  return(growthDf)
}

for (i in 0:3){
  print(i)
  }

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.