### 2021-06-07 R/Raw Img Data.Rmd

---
title: "Raw img Data"
author: "Juan P Barraza"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE}

#####
imgHisData<-read.csv("/Users/jupabago/Documents/Whiteley/PROJECTS/SaPa Chronic Wounds/Images data/Image histogram data/cowtd1m2p2a_1.csv",header = F)
NewNames <- c('zeroInt','zeroPercent','maxInt','firstZero','totalZeroes','level','calculatedG2','tailBins', 'numtailEnd','optimalG','meanIntensity','sdIntensity')
names(imgHisData) <- NewNames

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.