### 2021-06-29 R/PropOcDataThor.Rmd
Thor version of the Proportional occupancy code

---
title: "Proportional Occupancy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(gdata)
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(plotrix)
library(ggsignif)
library(rdist)

```

```{r functions to get data}
#using the new 3D script with capped spheres.
source('/home/jbarraza6/Documents/RScripts/PropOcFunctions2.0.R') 
GetPropOcData<-function(sample, timepoints, positions){
  #folder<- paste0('/media/jbarraza6/Seagate Backup Plus Drive/Good images/',sample,'/images1')
  folder<- paste0('~/Desktop/Seagate Backup Plus Drive/Good images/',sample,'/imagesSubtractRed1')
  growthDf<-data.frame()#data frame to add all timepoints and positions
  fileList<-list.files(path = folder, full.names = TRUE)
  print(folder)
  for (position in 2:positions){
    currentPosition<-fileList[grep(paste0("p",position,"_"),fileList)]
    print(paste0('analyzing position ', position))
    for (timepoint in 0:timepoints){
      currentTimePoint<-currentPosition[grep(paste0("t",GetSlice(timepoint),"_"),currentPosition)]
      dataTimePoint<-AllDataPipeline(SlicesTo3D(currentTimePoint,"red"),SlicesTo3D(currentTimePoint,"green"), 
                                     xydimSearch=114,zdimSearch = 70, sampleSize=1000,xyRealDim=0.264,
                                     zRealDim = 0.44, pipeBinSize = 1)
      dataTimePoint$TimePoint <- timepoint
      dataTimePoint$Position <- position
      growthDf<-rbind(growthDf,dataTimePoint)
      }
    }
  return(growthDf)
  }
GetSlice<-function(idx){
  if (idx<10){num = paste0(0,idx)}
  else
  {num = idx}
  return(num)}
```

```{r running functions}
wt1.1000<-GetPropOcData("3-13-19", 6, 2)
write.csv(wt1.1000, "~/Desktop/Seagate Backup Plus Drive/Good images/PropOcDataAll/wt1-1000-3.csv")

wt2.1000<-GetPropOcData("3-19-19", 6, 2)
write.csv(wt2.1000, "~/Desktop/Seagate Backup Plus Drive/Good images/PropOcDataAll/wt2-1000-3.csv")

wt3.1000<-GetPropOcData("4-17-19", 6, 2)
write.csv(wt3.1000, "~/Desktop/Seagate Backup Plus Drive/Good images/PropOcDataAll/wt3-1000-3.csv")

mut1.1000<-GetPropOcData("4-24-19", 6, 2)
write.csv(mut1.1000, "~/Desktop/Seagate Backup Plus Drive/Good images/PropOcDataAll/mut1-1000-3.csv")

mut2.1000<-GetPropOcData("4-25-19", 6, 2)
write.csv(mut2.1000, "~/Desktop/Seagate Backup Plus Drive/Good images/PropOcDataAll/mut2-1000-3.csv")
mut3.1000<-GetPropOcData("5-8-19", 6, 2)
write.csv(mut3.1000, "~/Desktop/Seagate Backup Plus Drive/Good images/PropOcDataAll/mut3-1000-3.csv")

```