### 2021-08-14 R/Prop Oc Data Clean.Rmd
new script looking only at prop Oc wt data

---
title: "Prop Oc Data Chronic Wounds"
author: "Juan P Barraza"
date: "8/14/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(gdata)
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(plotrix)
library(ggsignif)
library(rdist)
library(stringr)

lseq <- function(from, to, length.out) {
  # logarithmic spaced sequence
  # blatantly stolen from library("emdbook"), because need only this
  exp(seq(log(from), log(to), length.out = length.out))}

source('/Users/jupabago/Documents/Whiteley/PROJECTS/Rpackage/PropOc Package/R/PropOcFunctions2.0.R')

ReadPropOcFile<-function(file){
    propOcData<-read.csv(file,header = T)
    fileName<-str_split(basename(file), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
    print(paste0("Importing sample ", fileName))
    attrList<-str_split(fileName,"_")[[1]]
    propOcData<-propOcData %>% mutate(Condition = attrList[1], TimePoint = attrList[2], Replicate = attrList[3], SampleCode = fileName)
    return(propOcData)
    }

propOcDatafolder<-'/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/PropOc Data1'
propOcDatafileList<-list.files(path = propOcDatafolder, full.names = TRUE)
rawPropOcDataList<-map(propOcDatafileList,ReadPropOcFile)
rawPropOcData<-bind_rows(rawPropOcDataList)
#rawPropOcData<-rawPropOcData %>% filter(!SampleCode=="mono_d1_01Slim") %>%  filter(!SampleCode=="mono_d1_02Slim")

#QC for data:
##How much data is just straight up missing?
#####
countData<-rawPropOcData %>% group_by(SampleCode, source, Distance) %>% summarize(count = n()) %>% filter(count<1000) %>%ungroup() %>% summarise(MisingTotal = n()*1000-sum(count),TotalMedian = median(count), Mean = mean(count))

countData2<-rawPropOcData %>% group_by(SampleCode, source, Distance) %>% summarize(count = n()) %>% ungroup() %>%filter(count<1000) %>%ungroup() %>% group_by(SampleCode) %>% summarise(MisingTotal = n()*1000-sum(count),TotalMedian = median(count), Mean = mean(count), N = n()) %>% mutate(MissingPercent = MisingTotal/120000*100)

countDistancesALL<-rawPropOcData %>% group_by(SampleCode,source, Distance) %>% summarise(count = n()) %>% ungroup() %>%  mutate(difference = max(count)- count)
ggplot()+
  geom_point(data = countDistancesALL,aes(x=Distance,y=difference, shape = source, color = SampleCode),size=4)+
  labs(title ="Prop Oc Analysis Sampling missing by Source and Distance")+
  facet_grid(.~source)+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5))

#from 120 possible distance-source combinations, how many of those are missing? 
countDistancesBadOnly<- countDistancesALL%>% filter(difference>0) %>% group_by(SampleCode) %>%  summarise(missingDistantes = n())

#####
#How much of it is outside the ideal boundaries?
#####
CheckBenchmark<-function(benchmark){
#CheckPropOc<-rawPropOcData %>% filter(PropOc>benchmark|PropOc<0)
CheckPropOc<-rawPropOcData %>% filter(PropOc<0)
percentInbound<-length(CheckPropOc$SampleCode)/length(rawPropOcDataClean$SampleCode)*100
return(percentInbound)}

CheckBenchmark(1.5)
```